### 2014-01-30 paper/javascript.tex
Move paper to /paper.

\section{Example: JavaScript}
\label{sec:javascript}

(NB:\ the JS spec refers to execution context and state; when we refer to
execution context, we refer to both the context AND the state)

(Some background on JavaScript versus ECMAScript, since ECMAScript
constitutes the semantics we're actually going to combine with
IFC;\ everything else has to be locked down or made IFC-specific).

(Lay out how ECMAScript semantics work. Execution context and syntax.
Don't go into too much detail.  Point out that the execution context/state
corresponds implementation-wise to a JavaEngine)

(Minimal IFC fragment.  Actually, this wants to be in the general section.
So specialize this some more?)

(Combined ECMAScript and IFC\@. Talk about serialization. Draw parallel to web
workers, note the similarity.)