### 2014-02-14 paper/cesk.tex
conditions on label+fork; CSK -> CESK

\section{Generalizing to a CESK machine}
\label{sec:cesk}

Why is CSK not enough? Well if we want to write a correct condition
for label we need to make sure that if the target language is say just
STLC, then a lambda term that captures variables is properly handled.

(Definition of CESK)

Control-State-Environment-Kontinuations

** Insert definition from plt redex book **
We consider languages that can be modeled using imperatice SOS, with
an explicit store and evaluation contexts.

\paragraph{Conditions on label}
$\forall
\TE,
\Istore_1, \Il_1, \Ic_1,
\Istore_2, \Il_2, \Ic_2.
\Iconf{\Istore_1}{\Il_1}{\Ic_1}{\inTE{\Tv}} \Tarrow^* k
\Leftrightarrow              
\Iconf{\Tstore_2}{\Il_2}{\Ic_2}{\inTE{\Tv}} \Tarrow^* k
$
and\\
$
\Iconf{\Istore_1}{\Il_1}{\Ic_1}{\inTE{\Tv}} \Tarrow^*
\Iconf{\Istore'}{\Il'}{\Ic'}{\diverge}
\Leftrightarrow              
\Iconf{\Tstore_2}{\Il_2}{\Ic_2}{\inTE{\Tv}} \Tarrow^*
\Iconf{\Istore'}{\Il'}{\Ic'}{\diverge}
$

\begin{enumerate}
 \item What is this saying? In any evaluation context $\TE$ the value
 \Tv reduces to the same configuration regardless of the store (and
 environment).
 \item We must consider the evaluation in the combined language; this
 e.g., allows labeling already-labeled values.
 \item The typing rule $\forall \Delta, \Gamma. \Delta;\Gamma \vdash
 \Tv : \tau$ is necessary when we want to preserve the type of \Tv,
 but no sufficient (e.g., can have same values of same type, but
 different value in the store; hence, we also need the SOS condition).
 \item The stronger condition: $\cdot \vdash \Tv : \tau$ implies the
 SOS condition.
\end{enumerate}

\paragraph{Conditions on fork}
If we do not allow sharing of state between threads then there is no
SOS condition for fork. We don't care if the term gets stuck in the
new thread because $\Tklone$ removed elements. However, we do need a
condition when we consider typed systems. The condition for this is
given in the typing rule for fork:
%\Ifork{\Il'}{(\IT{\Te})})
\begin{mathpar}
\inferrule
{
\Delta; \Gamma \vdash \Ie_1 : l\\
\Tklone_\Delta(\Delta); \Gamma \vdash \Ie_2 : \tau
}
{
\Delta; \Gamma \vdash \Ifork{\Ie_1}{\Ie_2} : \tau
}
\end{mathpar}