### 2014-02-20 paper/monad.tex
lhs makefile; put new outline
TODO: add notation for lhs stuff

\section{Monadic interpretation}
\label{sec:monad}