### 2014-02-26 paper/monad.tex
Start monads
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

\section{The monadic interpretation}
\label{sec:monad}

In this section, we discuss the intuition behind the operational rules
we have proposed by considering their interpretation when structured
using monads and monad transformers.  Monads were originally proposed
as a method for structuring operational semantics~\cite{Moggi:2007:SOS:1230168.1230592}