### 2014-04-26 paper/rebuttal.txt
Done
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

Reviewer 2 asks whether or not our method actually scales in practice,
pointing to our inability to deal with external effects as a mark
against it.  It is true that dealing with external effects is important
for pushing IFC into practice, and our work is not primarily concerned
with addressing this issue.  However, this problem is endemic to all
information flow control systems: for instance, JPMail [A] re-implements
primitives with external effects in Jif; Paragon [B] internalizes
external effects by wrapping the corresponding primitives to be
IFC-aware [C]; JSFlow [25] manually wraps JavaScript primitives with
external effects.  This is not to say that the systems we have
implemented do not deal with external effects: e.g. XMLHttpRequest as
used in SWAPI essential wraps the "native" unsafe code. However, the treatment
of external effects is very much application specific, and finding a
general solution is an open problem.

Thus, the generality of our system does not come from a treatment of
external effects; rather, assuming that external effects are taken care
of (Section 6), it comes from our ability to accommodate arbitrarily
complex internal semantics.  For example, our work is directly
applicable to "vanilla" JavaScript (ECMAScript), which does not have any
external effects.  As opposed to existing IFC systems (e.g., [25]) which incur
tremendous overhead, we show that you can build an efficient,
non-intrusive IFC system for JavaScript, or any other language which can
be isolated.

It's worth stating that we did not simply shoe-horn SWAPI and LIO into a
"unified framework".  While our language uses ideas from LIO and SWAPI,
we note that the current versions of
these systems coevolved with the concepts in the paper.  For instance,
in LIO [53] (and Breeze [28]) exceptions were originally treated in a
rather complex fashion. As we discuss in the paper, when you consider
exceptions as part of the internal semantics of the target language, the
appropriate semantics for exceptions fall out naturally in our framework.
In fact, LIO was modified to adopt this new, simpler approach.

Reviewer 1 was concerned that many aspects of our IFC system were simply
rehashes of prior work.  We believe that our core IFC language's support
for both concurrent and sequential scheduling through asynchronous
communication primitives is novel.  (Notice that a primitive which may
block cannot work in a sequential setting).  This is an old idea from IFC
OS work, but formally reasoning about this idea is new.

Reviewer 1 also asks whether or not the ideas in the extension section
were actually worked out.  The IFC language features and external effect
systems mentioned have been implemented for Haskell and JavaScript, and,
in the case of LIO, have had properties formally proven, while the
scheduling policies and type safety extensions are more speculative.
The treatment of mutable references was explicitly called out as
unworkable: we would like to clarify an important distinction between
mutable references in the IFC language and mutable references internally
supported in the target language.  The naive addition of IFC references
results in two *different* types of references: this is a feature,
not a bug, since it avoids the need of adding expensive checks to
existing references which, in all likelihood, are local to an IFC thread.

Other questions
---------------

As reviewer 3 points out, the reference about SWAPI does not include
an URL.  It is available at http://deian.net/swapi.pdf and currently
not indexed by search engines because the paper is being anonymously
submitted.  We apologize for the mistake.

Reviewer 1 ask why we have not developed our formalism with a monad.
One of our motivations was to formalize an IFC system for JavaScript,
whose semantics are typically not presented in a stratified way by
distinguishing a monadic and a pure fragment.  A non-monadic approach
enables a more natural formalism for a large class of languages,
including JavaScript.

Reviewer 2 ask about the proof strategy for Theorem 3.  Performing
induction on the length of the derivation sequence followed by
induction on the derivation tree is standard for small-step
operational semantics.  This is isomorphic to induction on the
structure of the derivation for ->*.

Reviewer 3 asks about shared mutable references:  It is correct that a
different design choice might allow shared mutable references in this
way.  However, because native references are protected by the current
label, this would effectively be a fine-grained IFC system, requiring
checks for all reference reads and writes.  This would be against our
goal of a coarse-grained system.  We will improve the unfortunate
phrasing.

Reviewer 3 asks about the values sent across channels in SWAPI.  For
simplicity, we require them to be strings in the paper, thought any
structurally cloned object (as well as labeled blobs) are actually
permitted.

Reviewer 1 ask about the soundness of the JS-send rule.  While omitted
in the paper, this rule can be proven to be non-interfering.
Intuitively, for an attacker at level L, the additional premise
evaluates regardless of secrets at a level above L, thus ensuring that
no secrets can be observed.

References
----------

[A] Boniface Hicks, Kiyan Ahmadizadeh, and Patrick McDaniel. Understanding
practical application development in security-typed languages. In 22nd Annual
Computer Security Applications Conference (ACSAC), Miami, Fl, December 2006.

[B] Paragon for Practical Programming with Information-Flow Control Niklas
Broberg, Bart van Delft and David Sands Asian Symposium on Programming Languages
and Systems (APLAS) 2013, Springer

[C] http://www.cse.chalmers.se/research/group/paragon/?p=Case%20Studies