### 2014-07-08 paper/rebuttal.txt
s/SWAPI/COWL
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

We would like to thank all reviewers for their feedback and suggestions.
We will first reply to high-level comments and then reply to specific
questions in the reviews.

Reviewer 2 asks whether or not our method actually scales in practice,
pointing to our inability to deal with external effects as a mark
against it.  It is true that dealing with external effects is
important for pushing IFC into practice, and our work is not primarily
concerned with addressing this issue.  However, this problem is
endemic to all information flow control systems: for instance, JPMail [A]
re-implements primitives with external effects in Jif; Paragon [B]
internalizes external effects by wrapping the corresponding primitives
to be IFC-aware [C]; JSFlow [25,26] manually wraps JavaScript
primitives with external effects.  This is not to say that the systems
we have implemented do not deal with external effects: e.g.
XMLHttpRequest as used in COWL essential wraps the "native" unsafe
code. However, the treatment of external effects is very much
application specific, and finding a general solution is an open
problem.

Thus, the generality of our system does not come from a treatment of
external effects; rather, assuming that external effects are taken
care of (Section 6), it comes from our ability to accommodate
arbitrarily complex internal semantics.  For example, our work is
directly applicable to "vanilla" JavaScript (ECMAScript), which does
not have any external effects.  As opposed to existing IFC systems
(e.g., [25]) which incur tremendous overhead, we show that you can
build an efficient, non-intrusive IFC system for JavaScript, or any
other language which can be isolated.

It's worth stating that we did not simply shoe-horn COWL and LIO into
a "unified framework".  While our language uses ideas from LIO and
COWL, we note that the current versions of these systems coevolved
with the concepts in the paper.  For instance, in LIO [53] (and Breeze
[28]) exceptions were originally treated in a rather complex fashion.
As we discuss in the paper, when you consider exceptions as part of
the internal semantics of the target language, the appropriate
semantics for exceptions fall out naturally in our framework.  In
fact, LIO was modified to adopt this new, simpler approach.

Reviewer 1 was concerned that many aspects of our IFC system were
simply rehashes of prior work.  We believe that our core IFC
language's support for both concurrent and sequential scheduling
through asynchronous communication primitives is novel.  (Notice that
a primitive which may block cannot work in a sequential setting).
This is an old idea from IFC OS work, but formally reasoning about
this idea is new.

Reviewer 1 also asks whether or not the ideas in the extension section
were actually worked out.  The IFC language features and external
effect systems mentioned have been implemented for Haskell and
JavaScript, and, in the case of LIO, have had properties formally
proven, while the scheduling policies and type safety extensions are
more speculative.  The treatment of mutable references was explicitly
called out as unworkable: we would like to clarify an important
distinction between mutable references in the IFC language and mutable
references internally supported in the target language.  The naive
addition of IFC references results in two *different* types of
references: this is a feature, not a bug, since it absolves us of the
responsibility of integrating the two semantics.  Furthermore,
existing code using old references does not usually need to
communicate over IFC boundaries, so code being written to take
advantage of the new IFC support can be written to explicitly use IFC
references.  This approach of adding features to the core IFC calculus
extends to our other proposed extensions, and their security proofs
are generalizable to our setting in the same way.

Specific questions
------------------

As reviewer 3 points out, the reference about COWL does not include
an URL.  It is available at http://deian.net/swapi.pdf and currently
not indexed by search engines because the paper is being anonymously
submitted.  We apologize for the mistake.

Reviewer 1 ask why we have not developed our formalism with a monad.
One of our motivations was to formalize an IFC system for JavaScript,
whose semantics are typically not presented in a stratified way by
distinguishing a monadic and a pure fragment.  A non-monadic approach
enables a more natural formalism for a large class of languages,
including JavaScript.

Reviewer 2 ask about the proof strategy for Theorem 3.  Performing
induction on the length of the derivation sequence followed by
induction on the derivation tree is standard for small-step
operational semantics.  This is isomorphic to induction on the
structure of the derivation for ->*.

Reviewer 3 asks about shared mutable references:  It is correct that a
different design choice might allow shared mutable references in this
way.  However, because native references are protected by the current
label, this would effectively be a fine-grained IFC system, requiring
checks for all reference reads and writes.  This would be against our
goal of a coarse-grained system.  We will improve the unfortunate
phrasing.

Reviewer 3 asks about the values sent across channels in COWL.  For
simplicity, we require them to be strings in the paper, thought any
structurally cloned object (as well as labeled blobs) are actually
permitted.

Reviewer 1 ask about the soundness of the JS-send rule.  While omitted
in the paper, this rule can be proven to be non-interfering.
Intuitively, for an attacker at level L, the additional premise
evaluates regardless of secrets at a level above L, thus ensuring that
no secrets can be observed.

References
----------

[A] Boniface Hicks, Kiyan Ahmadizadeh, and Patrick McDaniel. Understanding
practical application development in security-typed languages. In 22nd Annual
Computer Security Applications Conference (ACSAC), Miami, Fl, December 2006.

[B] Paragon for Practical Programming with Information-Flow Control Niklas
Broberg, Bart van Delft and David Sands Asian Symposium on Programming Languages
and Systems (APLAS) 2013, Springer

[C] http://www.cse.chalmers.se/research/group/paragon/?p=Case%20Studies