### 2014-10-16 paper/conclusion.tex
Pass conclusions

%\cut{
\section{Conclusion}
\label{sec:conclusion}

In this paper, we argued that when designing a coarse-grained IFC
system, it is better to start with a fully isolated, multi-task system
and work one's way back to the model of a single IFC-aware language.  We showed
a formal framework capable to prove secure (i.e. non-interferent) systems designed this way,
without needing to rely on details of the target language.
%, and
%we provided conditions on how to securely refine our formal semantics to
%consider optimizations required in practice.
We provide soundness guarantees for the COWL's JavaScript monitor by simply
connecting our framework to that implementation.
%% to two IFC implementations for JavaScript based on this formalism,
%% explained how our methodology improved an exiting IFC system for Haskell,
%% and proposed
%% an IFC system for C using hardware isolation.
By systematically applying ideas from IFC in operating systems to programming languages,
we hope to have elucidated some of the core design principles of coarse-grained,
dynamic IFC systems.

% In this paper, we develop a general coarse-grained approach to IFC
% based on dividing a system into relatively course computational units
% and tracking only communication between these isolated units.
% We give formal semantics for the core coarse-grained
% information flow control language and show how a large class of target languages
% can be combined with it to achieve provable
% non-interference.
% We then give a proof technique for showing  non-interference
% of a concrete semantics for a potentially optimized IFC language
% by means of an isomorphism, and identify a class of IFC language restrictions
% that preserves non-interference.
% We briefly describe ways to enrich the core IFC language with
% more advanced concepts such as labeled values, privileges, or a
% notation of clearance and
% connect our formal semantics to real implementations of
% coarse-grained IFC systems for Javascript, Haskell, and
% (in less detail) enforcing IFC in C.
% Leaving more detailed presentation to later work,
% we informally relate our results to a monadic interpretation
% of IFC which guided our design.
% }