### 2014-10-18 paper/real.tex

\section{Real World Languages}
\label{sec:real}

In this section, we describe two JavaScript IFC implementations
building on the theoretical foundations of this work.
%
Furthermore, we consider how our formalism could be applied to the C
programming language and connect it to a previous IFC system for
Haskell.
%
%% The goal of this section is to show the flexibility of the embedding in settings with
%% vastly different properties and discuss the
%% semantic gap that must be overcome to apply our formalism
%% to other systems.
%
%% Two of the systems we describe have been implemented: the Haskell
%% system~\cite{lio} and the COWL~\cite{swapi} system; we leave the
%% implementation of the IFC system for C to future work.

\subsection{JavaScript}
\label{sec:real:js}

JavaScript, as specified by
ECMAScript~\cite{ecma}, does not have any built-in
functionality for I/O.
%(we denote this language as |targetLangJS|).
%
For this language, which we denote by |targetLangJS|, the IFC system
|specLangJS roundrobinf| can be implemented by exposing IFC primitives
to JavaScript as part of the runtime, and running multiple instances
of the JavaScript virtual machine in separate OS-level threads.
%
Unfortunately, this becomes very costly when a system, such as a
server-side web application, relies on many tasks.
%
%% For instance, consider a server-side web application that
%% creates a new task for each HTTP request.  The overhead would
%% quickly become impractical.

Luckly, this issue is not unique to our work---browser layout engines
also rely isolating code of different origins (e.g., according to the
same-origin policy).
%
Since creating an OS thread for each iframe is expensive, both
the V8 and SpiderMonkey JavaScript engines provide a means for running
JavaScript code in isolation in a single OS thread,
on disjoint sub-heaps.
%
In V8, this unit of isolation is called a \emph{context}; in
SpiderMonkey, it is called a \emph{compartment}.
%
(We will use these terms interchangably.)
%
Each context is associated with a global object, which, by
default, implements the JavaScript standard library (e.g.,
\verb|Object|, \verb|Array|, etc.).
%
Naturally, we adopt contexts to implement our notion of tasks.

When JavaScript is embedded in browser layout engines, 
or in server-side platforms such as Node.js,
additional APIs such as the Document Object Model (DOM) or the file
system get exposed as part of the runtime system.
These features are exposed by extending the global object, just like
the standard library.  For this reason, it is easy to modify
these systems to remove arbitrary external effects when implementing
an IFC system so important effects can be reintroduced in a safe manner.

\begin{figure}[t]
\centerline{\includegraphics[width=0.7\columnwidth]{figs/node}}
\caption{\label{fig:node}
This example shows how our trusted monitor (left) is used to mediate
communication between two tasks for which IFC is enforced (right).}
\end{figure}
%
\subsubsection{Server-side Node.js IFC:}\ 
We have implemented |specLangJS seqf| for Node.js in the form of a
library, without modifying Node.js or the V8 JavaScript engine.
%
Our implementation\footnote{Available at \codelink{}.} provides a
library for creating new tasks, i.e., contexts whose global object
only contains the standard JavaScript library and our IFC primitives
(e.g., |send| and |sandbox|).
%
When mapped to our formal treatement, |sandbox| is defined with |ap
klone tS = tS0|, where |tS0| is the global object corresponding to the
standard JavaScript library and our IFC primitives.
%
%These tasks can communicate and create new tasks, in turn.
%
These IFC operations are mediated by the trusted library code (executing
as the main Node.js context), which tracks the state (current label, messages,
etc.) of each task.  An example for |send|/|recv| is shown in
Figure~\ref{fig:node}.
%While V8 provides a means for sharing any object between contexts (used
%by same-origin contexts in Blink),
We remark that our system conservatively restricts
the kinds of messages that can be exchanged, via |send| (and |sandbox|),
to string values.
%
(We provide a convenience library which marshalls JSON objects to/from strings.)
%
In our formalization, this amounts to restricting the IFC language rule
for |send| in the following way:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  iv) , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' iv)), ldots)
.->
iS'; sched step (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
%

While the described system implements |specLangJS seqf|, applications
typically require access to libraries (e.g., the file system library
\textsf{fs}) that that have external effects.
%
Exposing the Node.js APIs directly to sandboxed tasks is unsafe.
Instead, we implement libraries such as a labeled \textsf{fs} as
message exchanges between the sandboxed tasks (e.g., \textsf{task-1}
in Figure~\ref{fig:node}) and the main Node.js task that implements
the IFC monitor.
%
While this is safer than simply wrapping unsafe objects, adding
featuers such as the \textsf{fs} requires the code in the main task to
ensures that labels are properly propagated and enforced.
%
Unfortunately, while imposing such a proof burden
is undesirable, this also has to be expected:
different language environments expose different libraries for
handling external I/O, and the correct treatment of external effects
is application specific.
%
%% For instance, in the case of Node.js, how labels interact with the
%% file system may even vary according to the application (e.g.,
%% in the IFC system Hails~\cite{hails}, labels are not persisted).
%
%While we can extend our formalism to model a particular interface to
%the file system, HTTP client, etc. the utility of this formalization is
%unclear and left to future work.
%
We do not extend our formalism to account for the  particular
interface to the file system, HTTP client, etc., as this is
specific to the Node.js implementation and does not generalize
to other systems.

\subsubsection{Client-side COWL IFC:}\
We now describe some of the implementation aspects of
COWL, a JavaScript implementation of coarse-grained
IFC for Mozilla's Gecko.  This work provides the formal basis for
the implementation, and COWL is described in much more detail in~\cite{swapi}.
%
Since the architecture of COWL is similar to that of Node.js, we focus
on the main differences.

%Tasks in COWL  correspond to browsing contexts, such as iframes,
%where the global object |tS0| also contains the DOM, the
%\verb|XMLHttpRequest| (XHR) object, etc.
%
Existing browser APIs offer \emph{web workers}, which can
be spawned by JavaScript code and have
the same isolation properties as tasks in our system.
%
The isolation of web workers was motivated by the lack of multithreading
in JavaScript; however, their isolation properties are equally useful for information
flow control.

To ensure that JavaScript code cannot communicate using the global object
(e.g., to the network with XHR or persistent storage using the DOM),
COWL additionally uses content security policy (CSP)~\cite{csp1.1}.\footnote{
  Since CSP does not yet restrict message-passing or navigation, this
  was addressed explicitly in the implementation.
}
%
Note, that this does not prevent tasks from manipulating the DOM
(e.g., to render information), so external effects due to DOM access
are also suppressed.
%
Simply disallowing all external effects for all browsing contexts is
overly-restricting; pages typically load images, perform XHR requests
to dynamically load data,
etc.
%
Hence, COWL allows
controlled network communication (inter-frame communication is
dictated by our |send| rule) by associating an implicit label with a
remote host: namely the label corresponding to its origin.
%
In turn, when a task perform a request, the COWL runtime performs the
same label check as that of |send|, namely that the task label can
flow to
the remote origin label.

Once again, while the external effects of COWL can be formally modeled,
we do not model these effects in our more general framework, since, like
for Node.js case, they are application specific.
%
%Indeed, even the labeled HTTP client in  Node.js and COWL differ, even
%though they're both JavaScript systems, simply because of the
%differing environments.

%\subsection{Haskell}
%\label{sec:real:hs}
%Our work borrows ideas from a Haskell coarse-grained IFC
%system called LIO~\cite{lio, stefan:addressing-covert}.
%%
%LIO relies on Haskell's strong type system and monadic encoding of
%side-effecting computations to achieve isolation and define the IFC
%sub-language.
%%
%Specifically LIO provides the \verb|LIO| monad as a way of restricting
%(almost all) side-effects; this monad also keeps track of a current
%label which is, in turn, used by trusted code when defining core IFC
%functions, such as |getLabel|.
%%
%In the context of our framework, LIO can be understood as follows: the
%\emph{pure subset} of Haskell is the target language, while the
%monadic subset of Haskell, operating in the \verb|LIO| monad, is the
%IFC language.
%
%%Indeed, by rephrasing LIO in terms of our framework, we simplified
%%LIO's treatment of exceptions.
%%
%LIO originally associated labels with exceptions, in a
%similar style to fine-grained
%systems~\cite{stefan:2012:arxiv-flexible, Hritcu:2013:YIB:2497621.2498098}.
%%
%In particular, the label of an exception was set to the current label
%at the time of the throw, and used by a custom catch construct to
%raise the current label, thereby preventing leaks due to exceptional
%control flows.
%%
%In addition to being overly complex, the interaction of exceptions
%with clearance (which sets an upper bound on the current label, see
%Section~\ref{sec:clearance}) was incorrect: the clearance
%was restored to the clearance at point of the catch (a consequence of
%defining \verb|LIO| using a State monad).
%%
%Furthermore, pure exceptions (e.g., divide by zero) always percolated to
%trusted code, effectively allowing for denial of service attacks.
%%
%The insights gained when viewing coarse-grained IFC as presented in this
%paper led to a much cleaner, simpler treatment of exceptions in LIO,
%which has now been adapted by LIO.
%%By using isolated tasks, and thus treating the current label (and
%%clearance) as global, the underlying Haskell exception system can be
%%used directly: exceptional control flow does not affect IFC.

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary system calls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
For instance, our notion of isolation can be achieved using Dune's
hardware protection mechanisms~\cite{Belay:2012:DSU:2387880.2387913},
similar to a simpler version of
Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611} with an information flow control
policy.
%
Using page tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates---tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept system
calls performed by
a task and only permit those corresponding to our IFC language (such as
|getLabel| or |send|).
%
Dune's hardware protection mechanism would allow us to provide a concrete
implementation that is efficient and relatively simple to reason
about, but other sandboxing mechanisms could be used in place of Dune.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a system call.
%
Creating a new task with the |sandbox| system call corresponds to
\emph{forking} a process.  Using page tables we can ensure that
there will be no shared memory
(effectively
defining |ap klone tS
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process.)
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these system calls are
meaningless.
%
We leave the implementation of this IFC system for C to future work.